---
title: "Midterm1Notes"
author: "Jasmine Hsiao"
date: "2024-02-05"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 1-"R, RStudio, and RMarkdown"
## Working Directory 
```{r}
getwd()
```

## Load the tidyverse
```{r}
library("tidyverse") 
```


# Lab 2.1-"Objects, Classes & NAs"
## Objects
```{r}
x <- 42 # the object 'x' 'gets' 42 
```

## Types of Data
Five frequently used `classes` of data: 1. numeric, 2. integer, 3. character, 4. logical, 5. complex.
```{r}
my_numeric <- 2
class(my_numeric) # To find the class of data
```

Use `is()` and `as()` functions to clarify or specify a type of data.
```{r}
is.integer(my_numeric) #is my_numeric an integer?
```

```{r}
my_integer <- as.integer(my_numeric) #create new object specified as an integer
```

## Missing Data
Functions to deal with NAs
```{r}
my_missing <- NA
```

```{r}
is.na(my_missing) # Asks if my_missing is NA
```

```{r}
anyNA(my_missing) # Asks if theres any NA
```

```{r}
new_vector <- c(10, 20, 30, NA)
mean(new_vector, na.rm=T) #removes the NA  in the vector. T = true, so it would be removed. 
```

# lab 2.2-"Vectors and Data Matrices"
## Vectors 
The `c` stands for concatenate, to create vectors
```{r}
my_vector <- c(10, 20, 30) # numeric vector
```

```{r}
days_of_the_week <- c("Monday", "Tuesday", "Wednesday", "Thrusday", "Friday", "Saturday", "Sunday") # character vector
```

```{r}
my_vector_sequence <- c(1:100) # generate sequence of numbers
my_vector_sequence
```

```{r}
my_vector_sequence[15] # Evalulate numeric in vector
```

## Data matrices
```{r}
Philosophers_Stone <- c(317.5, 657.1) # seperate vectors
Chamber_of_Secrets <- c(261.9, 616.9)
Prisoner_of_Azkaban <- c(249.5, 547.1)
Goblet_of_Fire <- c(290.0, 606.8)
Order_of_the_Phoenix <- c(292.0, 647.8)
Half_Blood_Prince <- c(301.9, 632.4)
Deathly_Hallows_1 <- c(295.9, 664.3)
Deathly_Hallows_2 <- c(381.0, 960.5)
```

```{r}
box_office <- c(Philosophers_Stone, Chamber_of_Secrets, Prisoner_of_Azkaban, Goblet_of_Fire, Order_of_the_Phoenix, Half_Blood_Prince, Deathly_Hallows_1, Deathly_Hallows_2)
box_office # New object called box_office
```

```{r}
harry_potter_matrix <- matrix(box_office, nrow = 8, byrow = T)
harry_potter_matrix # Creates a matrix
```

## Name the rows and columns
```{r}
region <- c("US", "non-US")
region
```

```{r}
titles <- c("Philosophers_Stone", "Chamber_of_Secrets", "Prisoner_of_Azkaban", "Goblet_of_Fire", "Order_of_the_Phoenix", "Half_Blood_Prince", "Deathly_Hallows_1", "Deathly_Hallows_2")
titles
```

```{r}
colnames(harry_potter_matrix) <- region # name column with vector 'region'
```

```{r}
rownames(harry_potter_matrix) <- titles # name rows with vector 'title'
```

```{r}
harry_potter_matrix # print matrix
```

## Using a data matrix
```{r}
global <- rowSums(harry_potter_matrix) # calculate total earnings
global
```

```{r}
all_harry_potter_matrix <- cbind(harry_potter_matrix, global)
all_harry_potter_matrix # Add new columns
```

```{r}
harry_potter_matrix[2,1] # selects the value in the first column, second row.
``` 

```{r}
harry_potter_matrix[1:4] #`:` selects the specified elements in a column. 
```

```{r}
non_us_earnings <- all_harry_potter_matrix[ ,2] #select values in an entire row or column, calculate mean of entire second column
mean(non_us_earnings)
```

# Lab 3.1-"Data Frames"
## Data Frames
Another type of data structure, similar to spreadsheet. Cleaner and neat looking than matrix.  
```{r}
Sex <- c("male", "female", "male") # Build seperate vectors
Length <- c(3.2, 3.7, 3.4)
Weight <- c(2.9, 4.0, 3.1)
```

```{r}
hbirds <- data.frame(Sex, Length, Weight) # Combine vectors to create data frame
hbirds
```

```{r}
names(hbirds) # Column names of data frame
```

```{r}
dim(hbirds) # Dimensions of hbirds
```

```{r}
str(hbirds) # Dimensions of hbirds
```

```{r}
hbirds <- data.frame(sex=Sex, length=Length, weight_g=Weight)
hbirds # Change names from uppercase to lowercase, When you say hbird gets data.frame, it overwrites everything
```

```{r}
hbirds[1,] # select first row
```

```{r}
hbirds[ ,3] # select third column
```

```{r}
mean(hbirds$weight_g) # select values in entire column using '$', useful for calculating means
```

## Adding Rows
```{r}
new_bird <- c("female", 3.6, 3.9)
new_bird
```

```{r}
hbirds<- rbind(hbirds, new_bird) # Add a new bird's data with 'rbind()' meaning row bind
hbirds
```

## Adding Columns
```{r}
hbirds$neighborhood <- c("lakewood", "brentwood", "lakewood", "scenic Heights")
hbirds # Add column into our data with '$'
```

## Writing Data to File
Save our hbirds data frame to use it again later. Here we write our data frame to a csv file. We use `row.names = FALSE` to avoid row numbers from printing out. 
```{r}
write.csv(hbirds, "hbirds_data.csv", row.names = FALSE)
```

# Lab 3.2-"Importing Data Frames"
## Load the tidyverse
```{r}
library("tidyverse")
```

```{r}
getwd() 
```

## Load the data
```{r}
hot_springs <- read_csv("lab3/hsprings_data.csv")
```

```{r}
str(hot_springs) # Get an idea of data structure
```

```{r}
class(hot_springs$scientist) 
class(hot_springs$spring) # shows the classes of both
```

```{r}
hot_springs$scientist <- as.factor(hot_springs$scientist)
hot_springs$spring <- as.factor(hot_springs$spring) # Makes both factors
```

```{r}
levels(hot_springs$scientist) 
levels(hot_springs$spring) # Shows levels of the factors
```

## Summary functions
```{r}
fish <- read_csv("lab3/data/Gaeta_etal_CLC_data.csv")
```

```{r}
summary(fish) # summarize data frame
```

```{r}
glimpse(fish) # Another summary function
```

```{r}
nrow(fish) # Gives number of rows
```

```{r}
ncol(fish) # Gives number of columns
```

```{r}
dim(fish) # Gives dimensions
```

```{r}
names(fish) #Gives column names
```

```{r}
head(fish) # Prints the first n rows of the data frame
```

```{r}
tail(fish) # Prints the last n rows of the data frame
```

```{r}
table(fish$lakeid) # Useful when have limited num of categorical variables, produces fast counts of number of observations in a variable
```

```{r}
# View(fish)
# click on the `fish` data frame in the Environment tab or type this
```

```{r}
little_fish <- filter(fish, length<=100)
little_fish # Filter allows us to pull out observations with specific criteria
```

# Lab 4.1-"Transforming data 1: Dplyr `select()`"
## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library("tidyverse")
```

## Load the data
```{r}
fish <- readr::read_csv("lab4/data/Gaeta_etal_CLC_data.csv") ##readr is just being explicit on the package used
```

## Data Structure
```{r}
glimpse(fish) # structure
summary(fish) # contents
dim(fish) # Dimensions
```