---
title: "Lab 6 Warmup"
author: "Jasmine Hsiao"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## 1. Load the `bison.csv` data.
```{r}
getwd()
```

```{r}
bison <- read_csv("data/bison.csv")
```
## 2. What are the dimesions and structure of the data?
```{r}
dim(bison)
```

```{r}
glimpse(bison)
```
## 3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
```{r}
names(bison)
```

```{r}
new_bison <- select(bison, "animal_code", "animal_sex", "animal_weight", "animal_yob")
#start using pipes instead
bison_new <- bison %>% 
  select(animal_code, animal_sex, animal_weight, animal_yob)
```

## 4. Pull out the animals born between 1980-1990.
```{r}
year_bison <- bison_new %>% 
  filter(between(animal_yob, 1980, 1990))
```
```{r}
bison_new %>% 
  filter(animal_yob>=1980 & animal_yob<=1990)
```


## 5. How many male and female bison are represented between 1980-1990?
```{r}
table(year_bison$animal_sex)
```
```{r}
males <- bison_new %>% 
  filter(animal_yob>=1980 & animal_yob<=1990)
  filter(animal_sex == "M")
males

table(males$animal_sex)
```
```{r}
females <- bison_new %>% 
  filter(animal_yob>=1980 & animal_yob<=1990)
  filter(animal_sex == "F")
females

table(females$animal_sex)
```

## 6. Between 1980-1990, were males or females larger on average?
Males are larger on average
```{r}
male_bison <- filter(year_bison, animal_sex == "M")
mean(male_bison$animal_weight, na.rm = T)
```

```{r}
female_bison <- filter(year_bison, animal_sex == "F")
mean(female_bison$animal_weight, na.rm = T)
```

```{r}
mean(males$animal_weight)
mean(females$animal_weight)
```

